---
title       : NMIS Mopup
subtitle    : the status, now, and future needs
author      : Data Team, Modi Research Group
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}

---

## Outline
1. Mopup -- a process overview
2. Outcomes of the matching process
3. Missing facilities
4. Baseline coverage, a retrospective

---

## Mopup, process overview
The stages of the mopup process include:

1. Facility Lists. Obtained by TAs, and entered using the formhub data tool to EI.
2. Data Matching. TAs matched the facilities submitted through baseline with data from baseline.
3. Analysis to scope out the work required for mopup (this presentation).
4. Teams go to field and survey the "outstanding" facilities.

---

## Facility List process overview

Facility lists were submitted for 773 of 774 LGAs. A quick visual comparison of the number of facilities per LGA, comparing the facility lists and the baseline list is belows (each point is an LGA).
Each point below the line is an LGA for which the number of facilities in the facility list was less than 80% of those in baseline; meaning the officials' lists are incomplete by a 20% margin.

```{r echo=F, warning=F, message=F, comment=NA, fig.height=4, fig.width=13, cache=TRUE}
  library(plyr)
  library(ggplot2)
  library(scales)
  # may need to change the following line to get running on your computer
  setwd("~/Code/presentations/prabhas/kigali-mopup-presentation/")
  matched <- read.csv("data/all_matched_data.csv")
  matched$ratio_fl_over_bl <- matched$total_facility_list / matched$total_baseline
  ggplot(data=matched, aes(x=total_baseline, y=total_facility_list)) + geom_point() + 
    geom_abline(slope=0.8) +
    labs(x="Total facilities from baseline", y="Total facilities from Facility Lists",
         title = "Comparison of baseline and facility lists; (slope of line = 0.8)")
```
There are  59 such lgas.

---

## Facility List Process Overview (2) 
The 59 lgas:
```{r echo=F, warning=F, message=F, comment=NA, fig.height=4, fig.width=13, cache=TRUE}
  shortfall_lgas <- arrange(subset(matched, total_facility_list/total_baseline < 0.8),
                      state, lga)
  lganamesdump = function(df1) {
    d_ply(df1, .(state), function (df) {
      print(strwrap(str_c(toupper(df[1,]$state), ": ", str_c(tolower(df$lga), collapse=", ")), 90))
      })
  }
  #lganamesdump(shortfall_lgas)

  library(rgeos)
  library(maptools)
  library(scales)
  library(classInt)
  lgashp <- readShapeSpatial('~/Dropbox/Nigeria/Nigeria 661 Baseline Data Cleaning/raw_data/nga_lgas/nga_lgas.shp')
  lgas <- fortify(lgashp, region="lga_id")
  lga_map <- function(geom_map_thingy) {
    ggplot() + geom_map_thingy + expand_limits(x=lgas$long, y=lgas$lat) +
      theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "bottom")
  }
  matched$RatioCategory = cut(matched$ratio_fl_over_bl, breaks=c(0,.8,1.2,2.0,29), right=F)
  levels(matched$RatioCategory) <- c("0-80% (small facility lists)", "80-120% (roughly the same)", "120-200% (up to double)", "200% + (more than double)")
  p <- lga_map(geom_map(data = matched, aes(map_id = lga_id, fill = RatioCategory), map=lgas)) 
```
```{r echo=F, warning=F, message=F, comment=NA, fig.height=8, fig.width=13, cache=TRUE}
  p + scale_fill_manual(values = c("#F4A582", "#F7F7F7", "#92C5DE", "#0571B0")) # colorbrewer, turned 5-scale into 4-scale
```